
---
title: "Lab 5 - Hypothesis Testing"
author: "Your name"
date: "Date of lab session"
output: 
  html_document: 
    highlight: tango
    theme: spacelab
---

```{r setup, include = FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

* * *

## Lab report

**Load data**
```{r load_data, message = FALSE}
curry_data <- read_csv("https://dyurovsky.github.io/85309/data/lab5/curry_data.csv")
```


#### Exercise 1: 
```{r}
mean_shooting_prob <- curry_data %>%
  summarise(mean = mean(SHOT_MADE)) %>%
  pull()

sd_shooting_prob <- curry_data %>%
  summarise(mean = sd(SHOT_MADE)) %>%
  pull()

by_game_data <- curry_data %>%
  group_by(GAME_ID) %>%
  summarise(shooting_prob = mean(SHOT_MADE),
            sd = sd(SHOT_MADE))

ggplot(by_game_data, aes(x = shooting_prob)) +
  geom_histogram(bins = 15) +
  geom_vline(aes(xintercept = mean_shooting_prob), color = "gold", size = 2)

```
The distribution looks roughly normal and symmetric. The mean is around .5, and the the spread is pretty broad--there look to be games where Steph makes around 60-70% of his shots, and also games where he makes 30%

#### Exercise 2:
H0: percent of shots made after making a shot is not different from the percent of shots made after missing a shot
HA: percent of shots made after making a shot is different from the percent of shots made after missing a shot

#### Exercise 3:
I don't want the first shot of a game to count as coming after the last shot of the previous game--I want to throw it out from my analysis and not count is as a shot after a missed shot or a made shot

#### Exercise 4:
```{r}
lag_data <- curry_data %>%
  group_by(GAME_ID) %>%
  mutate(lag_shot = lag(SHOT_MADE))

eda_hothands <- lag_data %>%
  group_by(lag_shot) %>%
  summarise(shooting_prob = mean(SHOT_MADE))

eda_hothands
```
This data doesn't look consistent with the hot hands hypothesis because Steph is *less* likely to make a shot after he made his last shot relative to missing his last shot. The NA values tell me about the first shot in the game that Steph takes.

#### Exercise 5:
```{r}
# Number of shots taken after shots that were made
hot_shots <- lag_data %>%
  filter(lag_shot) %>%
  nrow() # count the number of rows

# Number of shots made after shots that were made
hot_made <- lag_data %>%
  filter(lag_shot & SHOT_MADE) %>%
  nrow()

# Number of shots taken after shots that were missed
not_shots <- lag_data %>%
  filter(!lag_shot) %>%
  nrow()

# Number of shots made after shots that were missed
not_made <- lag_data %>%
  filter(!lag_shot & SHOT_MADE) %>%
  nrow()

simulate_null <- function() {

  # Make a list with the right number of shots of each type
  shots <- c(rep("Hot", hot_shots), rep("Not", not_shots))
  
  # randomly select the made shots from this list
  made <- sample(shots, hot_made + not_made) 
  
  # Compute the difference shot success between hot and not shots
  random_hot_made <- sum(made == "Hot") / hot_shots
  random_not_made <- sum(made == "Not") / not_shots
  
  random_hot_made - random_not_made
}

simulate_null()

null_diffs <- tibble(sample = 1:1000, 
                     diff = replicate(1000, simulate_null()))


empirical_diff <- (hot_made / hot_shots) - (not_made / not_shots)
empirical_diff

ggplot(null_diffs, aes(x = diff)) + 
  geom_histogram() +
  geom_vline(aes(xintercept = empirical_diff), color = "gold", size = 2)
```
It's hard to tell exactly from the plot, but it looks like Steph might be reliably worse after he made his last shot. So maybe some evidence for cold hands.


### More practice:

#### Exercise 6:
```{r}
mean(empirical_diff > pull(null_diffs, diff))
```
The empirical data fall at around the 2nd percentile, which suggests that we should reject the null hypothesis (because the data are in the bottom 2.5% percent of the null distribution). But it's pretty borderline.





#### Exercise 7:

```{r}
# enter your code for Exercise 7 here
```

#### Exercise 8:

```{r}
# enter your code for Exercise 8 here
```

#### Exercise 9:
